---
title: "Session 1"
author: "Ivon Teresa Sánchez Cárdenas (ivonsanchez.01@estudiant.upf.edu)"
date: "Last update: `r format(Sys.time(), '%d %B, %Y')`"      
output:
  html_document:
    toc: true
    fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Analysis of the Heart Disease Dataset 
Load the data from
[here](https://raw.githubusercontent.com/jpinero/DMI_2021/main/datasets/heart_disease_dataset.csv), and the description is [here](https://raw.githubusercontent.com/jpinero/DMI_2021/main/datasets/heart_disease_description.txt). 
The original dataset comes from [here](https://archive.ics.uci.edu/ml/datasets/Heart+Disease) and corresponds to the [processed cleveland data](https://archive.ics.uci.edu/ml/machine-learning-databases/heart-disease/processed.cleveland.data)

## Perform an EDA on the dataset

```{r}

data<-read.csv(file = "heart_disease_dataset.csv", header = T, sep=" ", na.strings = c("NA", " ", "-999", "NULL","?"))
```


```{r}
str(data)
head(data)
summary(data)
View(data)
```
## Here we are handling missing values
```{r}
#DATA CLEANING

# find location of missing values
print("Position of missing values ")
which(is.na(data))
 
# count total missing values 
print("Count of total missing values ")
sum(is.na(data))

data <- na.omit(data)
print("Count of total missing values ")
sum(is.na(data))

```

```{r}
# separate variables in continuous or categorical
data$ca <- as.numeric(data$ca)
data$thal <- as.numeric(data$thal)

categorical_v <- c("sex", "cp", "fbs", "restecg", "exang", "slope", "ca", "thal", "num")

continuous_v <- names(data)[!(names(data) %in% c(categorical_v, "patient_id"))]

```
## Now we will analyse catagorical data (UNIVARIATE ANALYSIS)

```{r}
#sex variable
#data$sex_factor<-factor(data$sex)
#levels(data$sex_factor) <- c("F", "M")
#subset_sex<-data.frame(data$sex_factor)

cat_plots <- lapply(categorical_v, function(var) {
  ggplot(data, aes_string(var)) +
    geom_bar(fill = "blue") +
    theme_minimal() +
    ggtitle(paste("Distribution of", var))
})
grid.arrange(grobs = cat_plots, ncol = 3)
```


```{r}
# chest pain type
# : chest pain type
#        -- Value 1: typical angina
#        -- Value 2: atypical angina
#        -- Value 3: non-anginal pain
#        -- Value 4: asymptomatic
data$cp_factor <-factor(data$cp)
levels(data$cp_factor) <-c("typical angina","atypical angina", "non-anginal pain", "asymptomatic"  )
levels(data$cp_factor)
subset_cp<-data.frame(data$cp_factor)
barplot(table(subset_cp))
```

```{r}
library(ggplot2)
library(gridExtra)
library(corrplot)


```
#exercise induced angina
#exercise induced angina (1 = yes; 0 = no)
data$exang

#slope: the slope of the peak exercise ST segment
#        -- Value 1: upsloping
#        -- Value 2: flat
#        -- Value 3: downsloping
data$slope

#thal 3 = normal; 6 = fixed defect; 7 = reversable defect
data$thal

#diagnosis of heart disease (angiographic disease status)
#        -- Value 0: < 50% diameter narrowing
#        -- Value 1: > 50% diameter narrowing
data$num
```

## Create visualizations in order to show which variables seem to be more associated with heart disease

```{r}

```


# 2 Difference in mortality rates in hospitalized COVID-19 patients 
Using the supplementary material from the [Difference in mortality rates in hospitalized COVID-19 patients identified by cytokine profile clustering using a machine learning approach: An outcome prediction alternative](https://www.frontiersin.org/articles/10.3389/fmed.2022.987182/full), perform the following tasks

## Reproduce Figure 1 from the publication

```{r}

```


## Reproduce Figure 2 from the publication
but instead of representing the clusters in the annotation, represent the groups (G1 to G4)

```{r}

```

## Improve figure 2 of the publication
Add a second annotation with information of deathm and a third one with information of gender

```{r}

```


# session info {.unnumbered}

```{r, results='asis',  echo=FALSE, message=FALSE }
sessionInfo()
```
